---
title: "Whole Genome NIPT Test Result"
header-includes:
- \usepackage{fancyhdr}
- \fancypagestyle{plain}{\pagestyle{fancy}}
- \renewcommand{\headrulewidth}{0pt}
- \fancyhf{}
- \pagestyle{fancy}
- \fancyfoot[CO,CE]{NIPT Report version 4.1 (bashable, October 2025)}
- \fancyfoot[LE,RO]{\thepage}
geometry: margin=1.5cm
output:
  pdf_document:
    highlight: zenburn
    keep_tex: yes
    latex_engine: xelatex
    fig_caption: no
classoption: portrait
#mainfont: Calibri
---

```{r setup, include=FALSE}
knitr::opts_chunk$set(echo = FALSE, fig.pos = 'H')
```

```{r libs, results='hide', message=FALSE, warning=FALSE}
library(ggplot2)
library(gridExtra)
library(knitr)
library(kableExtra)
suppressPackageStartupMessages(library(dplyr))
library(tinytex)
```

```{r CONFIG - EDIT ME}
# ========= EDIT THESE PATHS / IDs =========
#sample_id <- "SAMPLE_001_ABC"  # <--- EDIT ME

# --- CONFIG (edit paths if needed) ---

# sample_id from CLI
args <- commandArgs(trailingOnly = TRUE)
sample_id <- if (length(args) >= 1) args[[1]] else "UNKNOWN"
if (sample_id == "UNKNOWN") stop("Usage: Rscript -e \"rmarkdown::render('reportgeneration.Rmd')\" --args <sample_id>")

# RDS datasets
nipt_dataset_path                 <- "analysisScripts/plotData/nipt_dataset_744_samples_newref716.rds"
nipt_trisomy_mix_dataset_25M_path <- "analysisScripts/plotData/nipt_trisomy_mix_dataset_25M_newref716.rds"
nipt_trisomy_mix_dataset_50M_path <- "analysisScripts/plotData/nipt_trisomy_mix_dataset_50M_newref716.rds"
nipt_trisomy_mix_dataset_75M_path <- "analysisScripts/plotData/nipt_trisomy_mix_dataset_75M_newref716.rds"
nipt_trisomy_mix_dataset_100M_path<- "analysisScripts/plotData/nipt_trisomy_mix_dataset_100M_newref716.rds"

# Per-sample inputs
results_dir <- file.path("results", sample_id, "intermediateOutput")

seqff_file                        <- file.path(results_dir, paste0(sample_id, ".seqff.txt"))
bowtie2_statistics_file           <- file.path(results_dir, paste0(sample_id, ".wisecondorx-b37_statistics.txt"))
wisecondorx_gender_file           <- file.path(results_dir, paste0(sample_id, ".wisecondorx-b37_gender.txt"))
wisecondor_filtered_idxstats_file <- file.path(results_dir, paste0(sample_id, ".bowtie2-b37-idxstats.txt"))
wisecondor_report_file            <- file.path(results_dir, paste0(sample_id, ".wisecondor-bowtie2_b37-report.txt"))
wisecondor_pdf_file               <- file.path(results_dir, paste0(sample_id, "-wisecondor-bowtie2_b37_z.pdf"))
wisecondorx_png_file              <- file.path(results_dir, paste0(sample_id, ".wisecondorx-b37.plots/genome_wide.png"))
wisecondor_cwz_csv_file           <- file.path(results_dir, paste0(sample_id, ".wisecondor-bowtie2_b37_cwz.csv"))
wisecondorx_bed_file              <- file.path(results_dir, paste0(sample_id, ".wisecondorx-b37_aberrations.bed"))



# ---- PATH SANITY CHECK (runs before any read.table/readRDS) ----
paths <- c(
  seqff_file,
  bowtie2_statistics_file,
  wisecondorx_gender_file,
  wisecondor_filtered_idxstats_file,
  wisecondor_report_file,
  wisecondor_pdf_file,
  wisecondor_cwz_csv_file,
  nipt_dataset_path,
  nipt_trisomy_mix_dataset_25M_path,
  nipt_trisomy_mix_dataset_50M_path,
  nipt_trisomy_mix_dataset_75M_path,
  nipt_trisomy_mix_dataset_100M_path
)

#cat("Knit working directory:\n  ", normalizePath(getwd(), winslash = "/"), "\n\n")
#for (p in paths) {
#  cat("CHECK:\n  ", p, "\n  -> ", normalizePath(p, winslash = "/", mustWork = FALSE),
#      "\n  exists: ", file.exists(p), "\n\n", sep = "")
#}

missing <- paths[!file.exists(paths)]
if (length(missing)) {
  stop("Missing input(s):\n", paste0(" - ", missing, collapse = "\n"),
       "\n\nCheck working directory and relative paths.")
}

 stopifnot(file.exists(wisecondorx_png_file))





# =========================================
```

```{r LOAD}
sample_id_short <- strsplit(sample_id, "_")[[1]][1]

# Load data
nipt_dataset <- readRDS(nipt_dataset_path)
nipt_dataset$seqff_fetal_fraction <- nipt_dataset$seqff_fetal_fraction * 100
nipt_dataset <- nipt_dataset[nipt_dataset$seqff_fetal_fraction > 0.0,]

nipt_trisomy_mix_dataset_25M <- readRDS(nipt_trisomy_mix_dataset_25M_path)
nipt_trisomy_mix_dataset_25M_T13 <- nipt_trisomy_mix_dataset_25M[startsWith(basename(nipt_trisomy_mix_dataset_25M$Sample), "T13"),]
nipt_trisomy_mix_dataset_25M_T18 <- nipt_trisomy_mix_dataset_25M[startsWith(basename(nipt_trisomy_mix_dataset_25M$Sample), "T18"),]
nipt_trisomy_mix_dataset_25M_T21 <- nipt_trisomy_mix_dataset_25M[startsWith(basename(nipt_trisomy_mix_dataset_25M$Sample), "T21"),]

nipt_trisomy_mix_dataset_50M <- readRDS(nipt_trisomy_mix_dataset_50M_path)
nipt_trisomy_mix_dataset_50M_T13 <- nipt_trisomy_mix_dataset_50M[startsWith(basename(nipt_trisomy_mix_dataset_50M$Sample), "T13"),]
nipt_trisomy_mix_dataset_50M_T18 <- nipt_trisomy_mix_dataset_50M[startsWith(basename(nipt_trisomy_mix_dataset_50M$Sample), "T18"),]
nipt_trisomy_mix_dataset_50M_T21 <- nipt_trisomy_mix_dataset_50M[startsWith(basename(nipt_trisomy_mix_dataset_50M$Sample), "T21"),]

nipt_trisomy_mix_dataset_75M <- readRDS(nipt_trisomy_mix_dataset_75M_path)
nipt_trisomy_mix_dataset_75M_T13 <- nipt_trisomy_mix_dataset_75M[startsWith(basename(nipt_trisomy_mix_dataset_75M$Sample), "T13"),]
nipt_trisomy_mix_dataset_75M_T18 <- nipt_trisomy_mix_dataset_75M[startsWith(basename(nipt_trisomy_mix_dataset_75M$Sample), "T18"),]
nipt_trisomy_mix_dataset_75M_T21 <- nipt_trisomy_mix_dataset_75M[startsWith(basename(nipt_trisomy_mix_dataset_75M$Sample), "T21"),]

nipt_trisomy_mix_dataset_100M <- readRDS(nipt_trisomy_mix_dataset_100M_path)
nipt_trisomy_mix_dataset_100M_T13 <- nipt_trisomy_mix_dataset_100M[startsWith(basename(nipt_trisomy_mix_dataset_100M$Sample), "T13"),]
nipt_trisomy_mix_dataset_100M_T18 <- nipt_trisomy_mix_dataset_100M[startsWith(basename(nipt_trisomy_mix_dataset_100M$Sample), "T18"),]
nipt_trisomy_mix_dataset_100M_T21 <- nipt_trisomy_mix_dataset_100M[startsWith(basename(nipt_trisomy_mix_dataset_100M$Sample), "T21"),]

bowtie2_statistics <- read.table(bowtie2_statistics_file, fill=TRUE, sep="\t", header=TRUE, nrows=23)
bowtie2_statistics_qc <- read.table(bowtie2_statistics_file, fill=TRUE, sep=":", skip=24)
wisecondorx_gender <- as.character(read.table(wisecondorx_gender_file)$V1)
wisecondor_filtered_idxstats <- read.table(wisecondor_filtered_idxstats_file, sep="\t", header=FALSE, nrows=25)
seqff <- read.table(seqff_file, sep=",", header=TRUE)

cwz_data <- data.frame(t(read.table(wisecondor_cwz_csv_file, sep=",", header=FALSE)$V1))
colnames(cwz_data) <- paste0("Chr", 1:22)
Chr13_z_score <- cwz_data$Chr13
Chr18_z_score <- cwz_data$Chr18
Chr21_z_score <- cwz_data$Chr21

wisecondor_report_arguments <- read.table(wisecondor_report_file, fill=TRUE, sep="=", header=FALSE, skip=1, nrows=3, strip.white=TRUE, stringsAsFactors=FALSE)
wisecondor_report_arguments_used_in_convert <- read.table(wisecondor_report_file, fill=TRUE, sep="=", header=FALSE, skip=5, nrows=6, strip.white=TRUE, stringsAsFactors=FALSE)
wisecondor_report_arguments_used_in_test <- read.table(wisecondor_report_file, fill=TRUE, sep="=", header=FALSE, skip=13, nrows=10, strip.white=TRUE, stringsAsFactors=FALSE)
wisecondor_report_bam_information <- read.table(wisecondor_report_file, fill=TRUE, sep="\t", header=FALSE, skip=25, nrows=5, strip.white=TRUE, stringsAsFactors=FALSE)
wisecondor_report_retro_filtering <- read.table(wisecondor_report_file, fill=TRUE, sep="\t", header=FALSE, skip=32, nrows=3, strip.white=TRUE, stringsAsFactors=FALSE)
wisecondor_report_z_score_checks <- read.table(wisecondor_report_file, fill=TRUE, sep="\t", header=FALSE, skip=36, nrows=3, strip.white=TRUE, stringsAsFactors=FALSE)
wisecondor_report_test_results <- read.table(wisecondor_report_file, fill=TRUE, sep="\t", header=TRUE, skip=41, strip.white=TRUE, stringsAsFactors=FALSE)
```

```{r PROCESS}
wisecondor_call_list <- data.frame(
  "Location" = wisecondor_report_test_results$location,
  "Size"     = wisecondor_report_test_results$mbsize,
  "Effect"   = wisecondor_report_test_results$effect,
  "Zscore"   = wisecondor_report_test_results$z.score,
  "Log2FC"   = round(log2(1 + wisecondor_report_test_results$effect/100),2),
  "FF"       = abs(wisecondor_report_test_results$effect)*2 
)

wisecondor_call_list <- wisecondor_call_list[wisecondor_call_list$Size > 0.1,]

if (nrow(wisecondor_call_list) > 0) {
  for (i in 1:nrow(wisecondor_call_list)) {
    if (wisecondor_call_list$Size[i] >= 10 & abs(wisecondor_call_list$Zscore[i]) >= 10 & wisecondor_call_list$Effect[i] >= 20) {
      wisecondor_call_list$Note[i] <- "LIKELY MATERNAL ABERRATION"
    } else if (wisecondor_call_list$Size[i] >= 10 & abs(wisecondor_call_list$Zscore[i]) >= 10 & wisecondor_call_list$Effect[i] < 20) {
      wisecondor_call_list$Note[i] <- "LIKELY FETAL ABERRATION"
    } else if (wisecondor_call_list$Size[i] < 10 & abs(wisecondor_call_list$Zscore[i]) >= 10) {
      wisecondor_call_list$Note[i] <- "QC not passed (size < 10 Mb)"
    } else if (wisecondor_call_list$Size[i] >= 10 & abs(wisecondor_call_list$Zscore[i]) < 10) {
      wisecondor_call_list$Note[i] <- "QC not passed (low z-score)"
    } else if (wisecondor_call_list$Size[i] < 10 & abs(wisecondor_call_list$Zscore[i]) < 10) {
      wisecondor_call_list$Note[i] <- "QC not passed (low z-score, size < 10 Mb)"
    }
  }
}

percentage_genomic_change <- round(sum(wisecondor_call_list$Size)/3000*100, digits=2)
number_of_calls <- nrow(wisecondor_call_list)
calls_passing_qc <- wisecondor_call_list[wisecondor_call_list$Size >= 10 & abs(wisecondor_call_list$Zscore) >= 10,]
number_of_calls_passing_qc <- nrow(calls_passing_qc) 

seqff_fetal_fraction <- seqff[1,2] * 100
seqff_fetal_fraction_rounded <- round(seqff_fetal_fraction, digits = 1)
seqff_fetal_fraction_rounded_percentage <- seqff_fetal_fraction_rounded

chrY_read_count <- wisecondor_filtered_idxstats[which(wisecondor_filtered_idxstats$V1=="Y"), 3]
chr1_read_count <- wisecondor_filtered_idxstats[which(wisecondor_filtered_idxstats$V1=="1"), 3]
chrY_vs_chr1_ratio <- ((chrY_read_count / chr1_read_count)-0.01)/2
chrX_median_logR <- bowtie2_statistics$ratio.median[which(bowtie2_statistics$chr=="X")]


current_sample_point <- data.frame(
  seqff_fetal_fraction = seqff_fetal_fraction,
  chrY_vs_chr1_ratio = chrY_vs_chr1_ratio,
  chrX_median_logR = chrX_median_logR,
  Chr13 = Chr13_z_score, Chr18 = Chr18_z_score, Chr21 = Chr21_z_score
)

reads_mapped   <- round(wisecondor_report_bam_information$V2[which(wisecondor_report_bam_information$V1=="Reads mapped:")] / 1e6, 2)
reads_unmapped <- round(wisecondor_report_bam_information$V2[which(wisecondor_report_bam_information$V1=="Reads unmapped:")] / 1e6, 2)
reads_duplicates <- round(wisecondor_report_bam_information$V2[which(wisecondor_report_bam_information$V1=="Reads rmdup:")] / 1e6, 2)
reads_lowqual <- round(wisecondor_report_bam_information$V2[which(wisecondor_report_bam_information$V1=="Reads lowqual:")] / 1e6, 2)
reads_passed <- reads_mapped - reads_lowqual
reads_total <- round((wisecondor_report_bam_information$V2[which(wisecondor_report_bam_information$V1=="Reads mapped:")] +
                      wisecondor_report_bam_information$V2[which(wisecondor_report_bam_information$V1=="Reads unmapped:")]) / 1e6, 2)

FF_cutoff <- if (reads_total >= 100.0) 2.0 else if (reads_total >= 50.0) 2.5 else 4.0

if (seqff_fetal_fraction >= 4.0 & reads_total >= 25.0) {
  qc_passed <- TRUE;  description_qc_result <- "QC PASSED"; description_fetal_fraction <- "OK"
  description_reads <- "OK. Minimum 25 mill. reads is required (FF >4.0%)"
} else if (seqff_fetal_fraction >= 4.0 & reads_total < 25.0) {
  qc_passed <- FALSE; description_qc_result <- "QC FAILED"; description_fetal_fraction <- "OK"
  description_reads <- "FAILED. Minimum 25 mill. reads is required (FF >4.0%)"
} else if (seqff_fetal_fraction >= 2.5 & reads_total >= 50.0) {
  qc_passed <- TRUE;  description_qc_result <- "QC PASSED"; description_fetal_fraction <- "OK"
  description_reads <- "OK. Minimum 50 mill. reads is required (FF 2.5% - 4.0%)"
} else if (seqff_fetal_fraction >= 2.5 & reads_total < 50.0) {
  qc_passed <- FALSE; description_qc_result <- "QC FAILED"; description_fetal_fraction <- "OK"
  description_reads <- "FAILED. Minimum 50 mill. reads is required (FF 2.5% - 4.0%)"
} else if (seqff_fetal_fraction >= 2.0 & reads_total >= 100.0) {
  qc_passed <- TRUE;  description_qc_result <- "QC PASSED"; description_fetal_fraction <- "OK"
  description_reads <- "OK. Minimum 100 mill. reads is required (FF 2.0% - 2.5%)"
} else if (seqff_fetal_fraction >= 2.0 & reads_total < 100.0) {
  qc_passed <- FALSE; description_qc_result <- "QC FAILED"; description_fetal_fraction <- "OK"
  description_reads <- "FAILED. Minimum 100 mill. reads is required (FF 2.0% - 2.5%)"
} else {
  qc_passed <- FALSE; description_qc_result <- "QC FAILED"
  description_fetal_fraction <- "FAILED. FF below limit"; description_reads <- ""
}

chrX_cutoff_max <- 0.016
chrX_cutoff_min <- -0.016
chrY_cutoff <- 0.001 # 0.0008
chrY_cutoff_max <- (0.058 * seqff_fetal_fraction / 100) + 0.002

if (chrY_vs_chr1_ratio > chrY_cutoff_max) {
  gender_xy <- "XYY"; description_gender_xy <- "ABNORMAL. Likely fetal sex chromosome abnormality."
} else if (chrY_vs_chr1_ratio < chrY_cutoff & chrX_median_logR > chrX_cutoff_min & chrX_median_logR < chrX_cutoff_max) {
  gender_xy <- "XX";  description_gender_xy <- "Female fetus."
} else if (chrY_vs_chr1_ratio >= chrY_cutoff & chrX_median_logR <= chrX_cutoff_min) {
  gender_xy <- "XY";  description_gender_xy <- "Male fetus."
} else if (chrY_vs_chr1_ratio < chrY_cutoff & chrX_median_logR <= chrX_cutoff_min) {
  gender_xy <- "X0";  description_gender_xy <- "ABNORMAL. Likely fetal sex chromosome abnormality."
} else if (chrY_vs_chr1_ratio < chrY_cutoff & chrX_median_logR >= chrX_cutoff_max) {
  gender_xy <- "XXX"; description_gender_xy <- "ABNORMAL. Likely fetal sex chromosome abnormality."
} else if (chrY_vs_chr1_ratio >= chrY_cutoff & chrX_median_logR > chrX_cutoff_min & chrX_median_logR < chrX_cutoff_max) {
  gender_xy <- "XXY"; description_gender_xy <- "ABNORMAL. Likely fetal sex chromosome abnormality."
} else if (chrY_vs_chr1_ratio >= chrY_cutoff & chrX_median_logR >= chrX_cutoff_max) {
  gender_xy <- "XXXY"; description_gender_xy <- "ABNORMAL. Likely fetal sex chromosome abnormality."
} else {
  gender_xy <- "Unknown"; description_gender_xy <- "Unable to determine fetal sex chromosome status."
}

description_chr13 <- "Normal chr13 result (Z-score between -10.0 and 10.0)"
description_chr18 <- "Normal chr18 result (Z-score between -7.0 and 7.0)"
description_chr21 <- "Normal chr21 result (Z-score between -7.0 and 7.0)"

if (Chr13_z_score >= 13.0) {
  description_chr13 <- "Trisomi 13 detected. High confidence (Z-score > 13.0)"
} else if (Chr13_z_score <= -13.0) {
  description_chr13 <- "Monosomy 13 detected. High confidence (Z-score < -13.0)"
} else if (Chr13_z_score >= 10.0) {
  description_chr13 <- "Likely trisomi 13 detected. WARNING: Weak Z-score (10.0 - 13.0)"
} else if (Chr13_z_score <= -10.0) {
  description_chr13 <- "Likely monosomy 13 detected. WARNING: Weak Z-score (between -13.0 and -10.0)"
}

if (Chr18_z_score >= 10.0) {
  description_chr18 <- "Trisomi 18 detected. High confidence (Z-score > 10.0)"
} else if (Chr18_z_score <= -10.0) {
  description_chr18 <- "Monosomy 18 detected. High confidence (Z-score < -10.0)"
} else if (Chr18_z_score >= 7.0) {
  description_chr18 <- "Likely trisomi 18 detected. WARNING: Weak Z-score (7.0 - 10.0)"
} else if (Chr18_z_score <= -7.0) {
  description_chr18 <- "Likely monosomy 18 detected. WARNING: Weak Z-score (between -10.0 and -7.0)"
}

if (Chr21_z_score >= 10.0) {
  description_chr21 <- "Trisomi 21 detected. High confidence (Z-score > 10.0)"
} else if (Chr21_z_score <= -10.0) {
  description_chr21 <- "Monosomy 21 detected. High confidence (Z-score < -10.0)"
} else if (Chr21_z_score >= 7.0) {
  description_chr21 <- "Likely trisomi 21 detected. WARNING: Weak Z-score (7.0 - 10.0)"
} else if (Chr21_z_score <= -7.0) {
  description_chr21 <- "Likely monosomy 21 detected. WARNING: Weak Z-score (between -10.0 and -7.0)"
}

descriptions_chromosomes <- c()
if(!grepl("Normal", description_chr13)) descriptions_chromosomes <- c(descriptions_chromosomes, description_chr13)
if(!grepl("Normal", description_chr18)) descriptions_chromosomes <- c(descriptions_chromosomes, description_chr18)
if(!grepl("Normal", description_chr21)) descriptions_chromosomes <- c(descriptions_chromosomes, description_chr21)

number_of_chromosome_calls <- length(descriptions_chromosomes)
if (number_of_chromosome_calls == 0) {
  descriptions_chromosomes <- "Normal result for chromosome 13, 18, and 21"
} else {
  descriptions_chromosomes <- paste(unique(descriptions_chromosomes), collapse = "; ")
}

if (qc_passed == TRUE) {
  if (number_of_calls_passing_qc == 0 & (gender_xy == "XY" | gender_xy == "XX")) { test_result <- "QC PASSED. NO ABERRATIONS FOUND" } 
  if (number_of_calls_passing_qc == 0 & gender_xy != "XY" & gender_xy != "XX") { test_result <- "QC PASSED. LIKELY FETAL SEX CHROMOSOME ABNORMALITY" } 
  if (number_of_calls_passing_qc > 0 & (gender_xy == "XY" | gender_xy == "XX")) { test_result <- paste0("QC PASSED. ", paste(unique(calls_passing_qc$Note), collapse = '; ')) }
  if (number_of_calls_passing_qc > 0 & gender_xy != "XY" & gender_xy != "XX") { test_result <- paste0("QC PASSED. ", paste(unique(calls_passing_qc$Note), ". LIKELY FETAL SEX CHROMOSOME ABNORMALITY", collapse = '; ')) }
} else {
  test_result <- "TEST QC FAILED"
}

item <- c("SampleID","Fetal fraction","Number of reads","Gender","Common aneuploidy calls","Variant calls (all)","Variant calls (QC filtered)")
value <- c(sample_id_short, paste0(seqff_fetal_fraction_rounded_percentage, " %"), paste0(reads_total, " million reads"), gender_xy, number_of_chromosome_calls, number_of_calls, number_of_calls_passing_qc)
description <- c("", description_fetal_fraction, description_reads, description_gender_xy, descriptions_chromosomes, "(Caution if more than 10 calls)", "(QC criteria: Size min. 10 Mb AND Z-score higher than +/-10)")
summary_table <- data.frame(value, description, row.names = item)
```

```{r sampleid_table, echo=FALSE, results='asis'}
kable(test_result, escape = TRUE, "latex", longtable = TRUE, booktabs = TRUE, align = "l", row.names = FALSE, col.names = NULL) %>%
  kable_styling(latex_options = c("striped","hold_position"), full_width = FALSE, font_size = 16)
```

## Summary
```{r summary_table, echo=FALSE, results='asis'}
kable(summary_table, "latex", longtable = TRUE, booktabs = TRUE, align = "l", row.names = TRUE, col.names = NULL) %>%
  kable_styling(latex_options = c("striped","hold_position"), full_width = FALSE, font_size = 10) %>%
  column_spec(1, width = "4cm") %>%
  column_spec(2, width = "3cm")
```

## Common aneuploidy calls

`r str(nipt_dataset)`
`r str(current_sample_point)`
`r str(Chr13_z_score)`
`r str(cwz_data)`

```{r chromosome_table, echo=FALSE, results='asis'}
chromosome_column_names <- c("Chromosome","Z-score","Description")
chromosome_names <- c("Chromosome 13","Chromosome 18","Chromosome 21")
chromosome_z_scores <- c(
  format(round(current_sample_point$Chr13, 2), nsmall = 2),
  format(round(current_sample_point$Chr18, 2), nsmall = 2),
  format(round(current_sample_point$Chr21, 2), nsmall = 2)
)
chromosome_descriptions <- c(description_chr13, description_chr18, description_chr21)
chromosome_table <- data.frame(chromosome_names, chromosome_z_scores, chromosome_descriptions)
colnames(chromosome_table) <- chromosome_column_names

kable(chromosome_table, "latex", longtable = TRUE, booktabs = TRUE, align = "l", row.names = FALSE) %>%
  kable_styling(latex_options = c("striped","hold_position"), full_width = FALSE, font_size = 10) %>%
  column_spec(1, width = "4cm") %>%
  column_spec(2, width = "2cm") %>%
  column_spec(3, width = "8cm")
```

## Variant calls (QC filtered)
```{r table_calls_passing_qc, echo=FALSE, results='asis'}
if (number_of_calls_passing_qc > 0) {
  kable(calls_passing_qc, "latex", longtable = TRUE, booktabs = TRUE, align = "l", row.names = FALSE) %>%
    kable_styling(latex_options = c("striped","hold_position"), full_width = FALSE, font_size = 10) %>%
    column_spec(1, width = "4.0cm") %>%
    column_spec(2, width = "1.3cm") %>%
    column_spec(3, width = "1.3cm") %>%
    column_spec(4, width = "1.3cm") %>%
    column_spec(5, width = "1.3cm") %>%
    column_spec(6, width = "1.3cm")
} else {
  cat("No variants called")
}
```


## Sex Chromosomes
```{r,echo=FALSE, warning=FALSE, out.width='.32\\linewidth', fig.width=3, fig.height=3}
ggplot(nipt_dataset, aes(x = seqff_fetal_fraction, y = chrX_median_logR)) + 
  scale_x_continuous(name="Fetal Fraction") + 
  scale_y_continuous(name="ChrX LogR (Median)") +
  geom_point(col = "darkgrey", size=0.3) +
  geom_hline(yintercept=chrX_cutoff_min, col = "green", linetype="dotted", size=0.6) +
  geom_hline(yintercept=chrX_cutoff_max, col = "green", linetype="dotted", size=0.6) +
  geom_vline(xintercept=FF_cutoff, col = "red", linetype="dotted", size=0.6) + 
  geom_point(data = current_sample_point, col = "blue", size=4) +
  ggtitle("ChrX vs. Fetal Fraction") +
  theme_bw()

ggplot(nipt_dataset, aes(x = seqff_fetal_fraction, y = chrY_vs_chr1_ratio)) + 
  scale_x_continuous(name="Fetal Fraction") + 
  scale_y_continuous(name="ChrY Ratio") +
  geom_point(col = "darkgrey", size=0.3) +
  geom_hline(yintercept=chrY_cutoff, col = "blue", linetype="dotted", size=0.6) +
  geom_vline(xintercept=FF_cutoff, col = "red", linetype="dotted", size=0.6) + 
  geom_point(data = current_sample_point, col = "blue", size=4) +
  ggtitle("ChrY vs. Fetal Fraction") +
  theme_bw() +
  geom_abline(intercept = 0.002, slope = 0.00058, col = "green", linetype = "dotted", size = 0.6)

ggplot(nipt_dataset, aes(x = chrX_median_logR, y = chrY_vs_chr1_ratio)) + 
  scale_x_continuous(name="ChrX LogR (Median)") + 
  scale_y_continuous(name="ChrY Ratio") +
  geom_point(col = "darkgrey", size=0.3) +
  geom_hline(yintercept=chrY_cutoff, col = "blue", linetype="dotted", size=0.6) +
  geom_vline(xintercept=chrX_cutoff_min, col = "green", linetype="dotted", size=0.6) +
  geom_vline(xintercept=chrX_cutoff_max, col = "green", linetype="dotted", size=0.6) +
  geom_point(data = current_sample_point, col = "blue", size=4) +
  ggtitle("ChrX vs. ChrY") +
  theme_bw()
```

\pagebreak


## Genome Overview (Wisecondor)
```{r}
include_graphics(wisecondor_pdf_file)
```

### Variant calls (all)
```{r call_list_table, echo=FALSE, results='asis'}
if (number_of_calls > 0) {
  kable(wisecondor_call_list, "latex", longtable = TRUE, booktabs = TRUE, align = "l", row.names = FALSE, caption = "All Wisecondor-called aberrations") %>%
    kable_styling(latex_options = c("striped","hold_position"), full_width = FALSE, font_size = 8) %>%
    column_spec(1, width = "4cm") %>%
    column_spec(2, width = "1.5cm") %>%
    column_spec(3, width = "1.5cm") %>%
    column_spec(4, width = "1.5cm")
} else {
  cat("No variants called")
}
```

\pagebreak

## Genome Overview (WisecondorX. XY fetuses show X-ratio drop according to fetal fraction)
```{r, echo=FALSE, message=FALSE, warning=FALSE, fig.align='center', out.width='90%', cache=FALSE}
png_abs <- normalizePath(wisecondorx_png_file, mustWork = FALSE)
stopifnot(file.exists(png_abs))
knitr::include_graphics(png_abs)










```{r wisecondorx_calls, echo=FALSE, results='asis', message=FALSE, warning=FALSE}
if (file.exists(wisecondorx_bed_file)) {
  wcx <- tryCatch(
    read.table(wisecondorx_bed_file, sep = "\t", header = TRUE, stringsAsFactors = FALSE, check.names = FALSE),
    error = function(e) read.table(wisecondorx_bed_file, sep = "\t", header = FALSE, stringsAsFactors = FALSE, check.names = FALSE)
  )

  # Ensure expected column names if missing
  expected <- c("chr","start","end","ratio","zscore","type")
  if (!all(expected %in% names(wcx))) {
    names(wcx)[seq_len(min(6, ncol(wcx)))] <- expected[seq_len(min(6, ncol(wcx)))]
  }



  # Coerce & clean
wcx$chr    <- gsub("^chr", "", as.character(wcx$chr))
wcx$start  <- as.numeric(wcx$start)
wcx$end    <- as.numeric(wcx$end)
wcx$ratio  <- as.numeric(wcx$ratio)
wcx$zscore <- as.numeric(wcx$zscore)

# Filter: hide chr X OR |z| < 10  (your OR requirement)
wcx <- wcx[!(wcx$chr == "X" | abs(wcx$zscore) < 10), , drop = FALSE]

# If nothing left, print a message and skip
if (nrow(wcx) == 0) {
  cat("No WisecondorX aberrations passing filters.")
} else {
  # Format positions (no scientific)
  start_str <- formatC(wcx$start, format = "f", digits = 0, big.mark = "")
  end_str   <- formatC(wcx$end,   format = "f", digits = 0, big.mark = "")

  # Derived columns
  wcx$Location <- paste0(wcx$chr, ":", start_str, "-", end_str)
  wcx$Size     <- round((wcx$end - wcx$start)/1e6, 2)
  wcx$Ldog2R   <- round(wcx$ratio, 4)
  wcx$Zscore   <- round(wcx$zscore, 2)
  wcx$Note     <- ifelse(wcx$Size > 10,
                         "QC passed (size > 10 Mb)",
                         "QC not passed (size < 10 Mb)")

  # Order and print
  chnum  <- suppressWarnings(as.integer(wcx$chr))
  ord    <- order(chnum, wcx$start, na.last = TRUE)
  wcx_out <- wcx[ord, c("Location","Size","Ldog2R","Zscore","type","Note")]

  kable(wcx_out, "latex", longtable = TRUE, booktabs = TRUE, align = "l", row.names = FALSE,
        caption = "All WisecondorX-called aberrations") %>%
    kable_styling(latex_options = c("striped","hold_position"), full_width = FALSE, font_size = 8)
}

	

}











```






\pagebreak

## Common Autosomal Chromosome Plots

### Chromosome 13
`r ifelse(grepl("Normal", description_chr13), description_chr13, paste0("\\textbf{\\textcolor{red}{", description_chr13, "}}"))`

```{r,echo=FALSE, message=FALSE, out.width='0.4\\linewidth', fig.width=6, fig.height=4.5}
get_label_position_T13 <- function(data, fraction = 0.75) {
  x_range <- range(data$seqff_fetal_fraction, na.rm = TRUE)
  x_pos <- x_range[1] + (x_range[2] - x_range[1]) * fraction
  model <- lm(Chr13 ~ seqff_fetal_fraction, data = data)
  y_pos <- predict(model, newdata = data.frame(seqff_fetal_fraction = x_pos))
  data.frame(seqff_fetal_fraction = x_pos, Chr13 = y_pos)
}
label_pos_T13_25M <- get_label_position_T13(nipt_trisomy_mix_dataset_25M_T13)
label_pos_T13_50M <- get_label_position_T13(nipt_trisomy_mix_dataset_50M_T13)
label_pos_T13_75M <- get_label_position_T13(nipt_trisomy_mix_dataset_75M_T13)
label_pos_T13_100M <- get_label_position_T13(nipt_trisomy_mix_dataset_100M_T13)

ggplot(nipt_dataset, aes(x = seqff_fetal_fraction, y = Chr13)) + 
  scale_x_continuous(name="Fetal Fraction") + 
  scale_y_continuous(name="Z-score") +
  geom_point(col = "gray50", size=0.8) +
  geom_hline(yintercept = 13, color = "blue", linetype = "dotted", size = 0.4) + 
  geom_hline(yintercept = 10, color = "grey", linetype = "dotted", size = 0.4) +
  geom_hline(yintercept = -13, color = "blue", linetype = "dotted", size = 0.4) + 
  geom_hline(yintercept = -10, color = "grey", linetype = "dotted", size = 0.4) +
  geom_vline(xintercept = FF_cutoff, col = "red", linetype="dotted", size=0.4) + 
  geom_point(data = current_sample_point, col = "blue", size=5) +
  ggtitle(paste0("Chr13 (Z-score: ", format(round(current_sample_point$Chr13, 2), nsmall = 2), ")")) +
  coord_cartesian(ylim = c(-20, 70)) +
  theme_bw() +
  theme(plot.title = element_text(size = 15)) +
  geom_smooth(data = nipt_trisomy_mix_dataset_25M_T13, aes(x = seqff_fetal_fraction, y = Chr13), method = "lm", color = "darkgrey", linetype = "dotted", size = 0.4, se = FALSE) +
  geom_smooth(data = nipt_trisomy_mix_dataset_50M_T13, aes(x = seqff_fetal_fraction, y = Chr13), method = "lm", color = "darkgrey", linetype = "dotted", size = 0.4, se = FALSE) +
  geom_smooth(data = nipt_trisomy_mix_dataset_75M_T13, aes(x = seqff_fetal_fraction, y = Chr13), method = "lm", color = "darkgrey", linetype = "dotted", size = 0.4, se = FALSE) +
  geom_smooth(data = nipt_trisomy_mix_dataset_100M_T13, aes(x = seqff_fetal_fraction, y = Chr13), method = "lm", color = "darkgrey", linetype = "dotted", size = 0.4, se = FALSE) +
  geom_text(data = label_pos_T13_25M, aes(label = "25M"), hjust = 0, vjust = -1, color = "darkgrey", size = 3) +
  geom_text(data = label_pos_T13_50M, aes(label = "50M"), hjust = 0, vjust = -1, color = "darkgrey", size = 3) +
  geom_text(data = label_pos_T13_75M, aes(label = "75M"), hjust = 0, vjust = -1, color = "darkgrey", size = 3) +
  geom_text(data = label_pos_T13_100M, aes(label = "100M"), hjust = 0, vjust = -1, color = "darkgrey", size = 3)
```

### Chromosome 18
`r ifelse(grepl("Normal", description_chr18), description_chr18, paste0("\\textbf{\\textcolor{red}{", description_chr18, "}}"))`

```{r,echo=FALSE, message=FALSE, out.width='0.4\\linewidth', fig.width=6, fig.height=4.5}
get_label_position_T18 <- function(data, fraction = 0.75) {
  x_range <- range(data$seqff_fetal_fraction, na.rm = TRUE)
  x_pos <- x_range[1] + (x_range[2] - x_range[1]) * fraction
  model <- lm(Chr18 ~ seqff_fetal_fraction, data = data)
  y_pos <- predict(model, newdata = data.frame(seqff_fetal_fraction = x_pos))
  data.frame(seqff_fetal_fraction = x_pos, Chr18 = y_pos)
}
label_pos_T18_25M <- get_label_position_T18(nipt_trisomy_mix_dataset_25M_T18)
label_pos_T18_50M <- get_label_position_T18(nipt_trisomy_mix_dataset_50M_T18)
label_pos_T18_75M <- get_label_position_T18(nipt_trisomy_mix_dataset_75M_T18)
label_pos_T18_100M <- get_label_position_T18(nipt_trisomy_mix_dataset_100M_T18)

ggplot(nipt_dataset, aes(x = seqff_fetal_fraction, y = Chr18)) + 
  scale_x_continuous(name="Fetal Fraction") + 
  scale_y_continuous(name="Z-score") +
  geom_point(col = "gray50", size=0.8) +
  geom_hline(yintercept = 10, color = "blue", linetype = "dotted", size = 0.4) + 
  geom_hline(yintercept = 7, color = "grey", linetype = "dotted", size = 0.4) +
  geom_hline(yintercept = -10, color = "blue", linetype = "dotted", size = 0.4) + 
  geom_hline(yintercept = -7, color = "grey", linetype = "dotted", size = 0.4) +
  geom_vline(xintercept = FF_cutoff, col = "red", linetype="dotted", size=0.4) + 
  geom_point(data = current_sample_point, col = "blue", size=5) +
  ggtitle(paste0("Chr18 (Z-score: ", format(round(current_sample_point$Chr18, 2), nsmall = 2), ")")) +
  coord_cartesian(ylim = c(-20, 70)) +
  theme_bw() +
  theme(plot.title = element_text(size = 15)) +
  geom_smooth(data = nipt_trisomy_mix_dataset_25M_T18, aes(x = seqff_fetal_fraction, y = Chr18), method = "lm", color = "darkgrey", linetype = "dotted", size = 0.4, se = FALSE) +
  geom_smooth(data = nipt_trisomy_mix_dataset_50M_T18, aes(x = seqff_fetal_fraction, y = Chr18), method = "lm", color = "darkgrey", linetype = "dotted", size = 0.4, se = FALSE) +
  geom_smooth(data = nipt_trisomy_mix_dataset_75M_T18, aes(x = seqff_fetal_fraction, y = Chr18), method = "lm", color = "darkgrey", linetype = "dotted", size = 0.4, se = FALSE) +
  geom_smooth(data = nipt_trisomy_mix_dataset_100M_T18, aes(x = seqff_fetal_fraction, y = Chr18), method = "lm", color = "darkgrey", linetype = "dotted", size = 0.4, se = FALSE) +
  geom_text(data = label_pos_T18_25M, aes(label = "25M"), hjust = 0, vjust = -1, color = "darkgrey", size = 3) +
  geom_text(data = label_pos_T18_50M, aes(label = "50M"), hjust = 0, vjust = -1, color = "darkgrey", size = 3) +
  geom_text(data = label_pos_T18_75M, aes(label = "75M"), hjust = 0, vjust = -1, color = "darkgrey", size = 3) +
  geom_text(data = label_pos_T18_100M, aes(label = "100M"), hjust = 0, vjust = -1, color = "darkgrey", size = 3)
```

### Chromosome 21
`r ifelse(grepl("Normal", description_chr21), description_chr21, paste0("\\textbf{\\textcolor{red}{", description_chr21, "}}"))`

```{r,echo=FALSE, message=FALSE, out.width='0.4\\linewidth', fig.width=6, fig.height=4.5}
get_label_position_T21 <- function(data, fraction = 0.75) {
  x_range <- range(data$seqff_fetal_fraction, na.rm = TRUE)
  x_pos <- x_range[1] + (x_range[2] - x_range[1]) * fraction
  model <- lm(Chr21 ~ seqff_fetal_fraction, data = data)
  y_pos <- predict(model, newdata = data.frame(seqff_fetal_fraction = x_pos))
  data.frame(seqff_fetal_fraction = x_pos, Chr21 = y_pos)
}
label_pos_T21_25M <- get_label_position_T21(nipt_trisomy_mix_dataset_25M_T21)
label_pos_T21_50M <- get_label_position_T21(nipt_trisomy_mix_dataset_50M_T21)
label_pos_T21_75M <- get_label_position_T21(nipt_trisomy_mix_dataset_75M_T21)
label_pos_T21_100M <- get_label_position_T21(nipt_trisomy_mix_dataset_100M_T21)

ggplot(nipt_dataset, aes(x = seqff_fetal_fraction, y = Chr21)) + 
  scale_x_continuous(name="Fetal Fraction") + 
  scale_y_continuous(name="Z-score") +
  geom_point(col = "gray50", size=0.8) +
  geom_hline(yintercept = 10, color = "blue", linetype = "dotted", size = 0.4) + 
  geom_hline(yintercept = 7, color = "grey", linetype = "dotted", size = 0.4) +
  geom_hline(yintercept = -10, color = "blue", linetype = "dotted", size = 0.4) + 
  geom_hline(yintercept = -7, color = "grey", linetype = "dotted", size = 0.4) +
  geom_vline(xintercept = FF_cutoff, col = "red", linetype="dotted", size=0.4) + 
  geom_point(data = current_sample_point, col = "blue", size=5) +
  ggtitle(paste0("Chr21 (Z-score: ", format(round(current_sample_point$Chr21, 2), nsmall = 2), ")")) +
  coord_cartesian(ylim = c(-20, 70)) +
  theme_bw() +
  theme(plot.title = element_text(size = 15)) +
  geom_smooth(data = nipt_trisomy_mix_dataset_25M_T21, aes(x = seqff_fetal_fraction, y = Chr21), method = "lm", color = "darkgrey", linetype = "dotted", size = 0.4, se = FALSE) +
  geom_smooth(data = nipt_trisomy_mix_dataset_50M_T21, aes(x = seqff_fetal_fraction, y = Chr21), method = "lm", color = "darkgrey", linetype = "dotted", size = 0.4, se = FALSE) +
  geom_smooth(data = nipt_trisomy_mix_dataset_75M_T21, aes(x = seqff_fetal_fraction, y = Chr21), method = "lm", color = "darkgrey", linetype = "dotted", size = 0.4, se = FALSE) +
  geom_smooth(data = nipt_trisomy_mix_dataset_100M_T21, aes(x = seqff_fetal_fraction, y = Chr21), method = "lm", color = "darkgrey", linetype = "dotted", size = 0.4, se = FALSE) +
  geom_text(data = label_pos_T21_25M, aes(label = "25M"), hjust = 0, vjust = -1, color = "darkgrey", size = 3) +
  geom_text(data = label_pos_T21_50M, aes(label = "50M"), hjust = 0, vjust = -1, color = "darkgrey", size = 3) +
  geom_text(data = label_pos_T21_75M, aes(label = "75M"), hjust = 0, vjust = -1, color = "darkgrey", size = 3) +
  geom_text(data = label_pos_T21_100M, aes(label = "100M"), hjust = 0, vjust = -1, color = "darkgrey", size = 3)
```

\pagebreak

## Sample details

Mapped reads: `r reads_mapped` mill. reads  
Unmapped reads: `r reads_unmapped` mill. reads  
Duplicates reads: `r reads_duplicates` mill. reads (`r as.character(round(reads_duplicates/reads_mapped*100, 1))`%)  
Low Quality reads: `r reads_lowqual` mill. reads  

ChrX logR ratio: `r as.character(round(chrX_median_logR, 5))`  
ChrY ratio: `r as.character(round(chrY_vs_chr1_ratio, 5))`  

Input sample_id: `r sample_id`  

**Wisecondor Report settings**  
mineffect: `r as.character(wisecondor_report_arguments$V2[which(wisecondor_report_arguments$V1=="mineffect")])`  
resultfile: `r wisecondor_report_arguments$V2[which(wisecondor_report_arguments$V1=="resultfile")]`  
testfile: `r wisecondor_report_arguments$V2[which(wisecondor_report_arguments$V1=="testfile")]`  

**Wisecondor arguments used in convert:**  
retthresh: `r wisecondor_report_arguments_used_in_convert$V2[which(wisecondor_report_arguments_used_in_convert$V1=="rethresh")]`  
retdist: `r wisecondor_report_arguments_used_in_convert$V2[which(wisecondor_report_arguments_used_in_convert$V1=="retdist")]`  
binsize: `r wisecondor_report_arguments_used_in_convert$V2[which(wisecondor_report_arguments_used_in_convert$V1=="binsize")]`  
outfile: `r wisecondor_report_arguments_used_in_convert$V2[which(wisecondor_report_arguments_used_in_convert$V1=="outfile")]`  
infile (for convert): `r wisecondor_report_arguments_used_in_convert$V2[which(wisecondor_report_arguments_used_in_convert$V1=="infile")]`  

**Wisecondor arguments used in test:**  
minzscore: `r wisecondor_report_arguments_used_in_test$V2[which(wisecondor_report_arguments_used_in_test$V1=="minzscore")]`  
minrefbins: `r wisecondor_report_arguments_used_in_test$V2[which(wisecondor_report_arguments_used_in_test$V1=="minretains")]`  
chromosomes: `r toString(wisecondor_report_arguments_used_in_test$V2[which(wisecondor_report_arguments_used_in_test$V1=="chromosomes")])`  
reference: `r wisecondor_report_arguments_used_in_test$V2[which(wisecondor_report_arguments_used_in_test$V1=="reference")]`  
outfile (for test): `r wisecondor_report_arguments_used_in_test$V2[which(wisecondor_report_arguments_used_in_test$V1=="outfile")]`  
mineffectsize: `r wisecondor_report_arguments_used_in_test$V2[which(wisecondor_report_arguments_used_in_test$V1=="mineffectsize")]`  
repeats: `r wisecondor_report_arguments_used_in_test$V2[which(wisecondor_report_arguments_used_in_test$V1=="repeats")]`  
infile (for test): `r wisecondor_report_arguments_used_in_test$V2[which(wisecondor_report_arguments_used_in_test$V1=="infile")]`  
multitest: `r wisecondor_report_arguments_used_in_test$V2[which(wisecondor_report_arguments_used_in_test$V1=="multitest")]`  

